Fully Extrinsic Type Inference (DRAFT 2025-06-20)

### 2025-09-04 doc/extrinsic_type_inference_paper/paper.tex
update checking interpretation in model typing
add notes on decidable check procedure
update subtyping check and tidy and interpret definitions in lean
update definitions to ensure completeness of negated predicates
update upper founded
define height and structural wrap
update struct less than
in paper, rename experiements to instances, since these are proved, not just externally observed

\newtheorem{experiment}{Experiment}[section]

@@

      \delta \satisfies e \hastype \J{EXI[}\Theta\J{]}\Delta \J{:} \tau
      &\iff& 
      \exists \delta' \qua
      \textbf{dom}(\delta') \subseteq \Theta
      \up
      \delta \oplus \delta' \satisfies \Delta 
      \up
      \delta \oplus \delta' \satisfies e \hastype \tau

@@

      \delta \satisfies e \hastype \J{ALL[}\Theta\J{]}\Delta \J{:} \tau
      &\iff& 
      \forall \delta' \qua
      \textbf{dom}(\delta') \subseteq \Theta
      \implies
      \delta \oplus \delta' \satisfies \Delta 
      \implies
      \delta \oplus \delta' \satisfies e \hastype \tau

@@

      \delta \satisfies e \hastype \alpha 
      &\iff& 
      \alpha \slash \tau \in \delta
      \up
      \VDash e \hastype \tau
  \end{array}
  \]
\end{definition}

@@

    \inferrule {
      \alpha \neq \alpha_l
      \\\\
      \textbf{invert}(\alpha \entails \Delta) = \Delta' 
      \\
      \textbf{pack}^-(\alpha\ \alpha_l\ \Theta^\dagger \entails \Theta, \Delta', \alpha_l \J{*} \tau_r) = \tau'
      \\\\
      \textbf{factor}(\alpha \entails \tau' \cdot \J{left}) = \tau_l
      \\
      \textbf{factor}(\alpha \entails \tau' \cdot \J{right}) = \tau_r'
      \\\\
      \alpha_l \wr^+ \tau_r'
      \\
      \textbf{found}(\alpha \entails \tau_l) = \tau_l'
      \\
      \tau_r'[\alpha_l\slash \J{LFP[} \alpha \J{]} \tau_l'] = \tau_r''
    } {
      \Theta^\dagger \entails \alpha \circlearrowleft (\Theta, \Delta, \alpha_l \J{->}\tau_r) \subtypes \alpha_l \J{->} \J{LFP[} \alpha \J{]} \tau_r''
    }
  \end{mathpar}
\end{definition}

@@

\begin{definition} 
  \label{def:proof_subtyping_difference_intro}
  \emph{proof subtyping (difference introduction)}
  \hfill
  \small
  \boxed{\Theta, \Delta \entails \tau \subtypes \tau_l \typdiff \tau_r \given \Theta', \Delta'}
  \nopad
  \begin{mathpar}
    \inferrule {
      \TODO{Update proof, deriving that lower does not intersect with subtraction}
      \\\\
      \TODO{in lean, prove that subtyping is equivalent to decidable definition under restricted subtyping form}
      \\\\
      \epsilon \pattype \tau_r
      \\
      \Theta, \Delta \restricted \tau_r \subtypes  \tau
      \\\\
      \Theta, \Delta \not\dentails \tau \subtypes \tau_r
      \\
      \Theta, \Delta \not\dentails \tau_r \subtypes \tau
      \\
      \Theta, \Delta \entails \tau \subtypes \tau_l \given \Theta', \Delta' 
    } {
      \Theta, \Delta \entails 
      \tau \subtypes \tau_l \typdiff \tau_r
      \given \Theta', \Delta'
    }
  \end{mathpar}
\end{definition}

@@

\section{Theory}
\label{sec:theory}

@@

\section{Experiments}
\label{sec:experiments}

\noindent
We performed experiments to see how type inference performs on both functions and constructors. 
These two categories have historically been tackled independently of each other, where parametric types
tend to rely on notions of instantiation, and structural types rely on notions of equality or subtyping.
Our system unifies these historically distinct problems by describing both kinds of types in terms of subtyping.
However, the distinction is useful for comparing our design to previous work.

We implemented an extrinsic type inference system that corresponds to the predicate definitions
presented thus far. 
To observe the expressiveness of our system on functions, 
we evaluated our implementation on the same suite of programs 
that SuperF was evaluated on, as presented in the paper on SuperF \cite{}.
Based on these experiments, our system all programs admitted by SuperF, as well as some programs 
that SuperF is unable to admit.
Thus, our system appears to be more expressive on functions than the previous most expressive
type inference system that we know of.
To observe the expressiveness of our system on constructors, 
we constructed simple programs that represent the kind of flexible and reusable code
often found programs written in untyped scripting languages. 
We ran these examples in our implementation, and created analogous version to run
in the online implementation of MLScript \cite{},
which represents the theoretical system MLStruct \cite{}. 
All of these examples are admitted by our system, while MLStruct
can either requires intrinsic specifications to admit the program, 
or simply can't admit the program.

\begin{definition} 
  \label{def:experimental_expressions_and_types}
  \emph{experimental expressions and types} 
  \scriptsize
  \nopad
  \[
  \begin{array}[t]{ll}
    \begin{array}[t]{r@{}c@{}l}
      \textbf{Bool} 
      &\ \ =\ \ & 
      \J{(<true>@ | <false>@)} 
      \\
      \textbf{Nat} 
      &\ \ =\ \ & 
      \J{(LFP[R] <zero>@ | <succ>R)} 
      \\
      \textbf{Even} 
      &\ \ =\ \ & 
      \J{(LFP[R] <zero>@ | <succ><succ>R)} 
      \\
      \textbf{List}(\alpha) 
      &\ \ =\ \ & 
      \J{(LFP[R] <nil>@ | <cons>((}\alpha\J{)*R))} 
      \\
      \textbf{ChurchNum} 
      &\ \ =\ \ & 
      \J{(ALL[T] (T -> T) -> T -> T)}
    \end{array}
    &
    \begin{array}[t]{r@{}c@{}l}
      \textbf{ST}(\alpha^\dagger, \alpha) 
      &\ \ =\ \ & 
      \J{(<tag>(}\alpha^\dagger \J{) <result>(} \alpha \J{))}
      \\
      \textbf{z} 
      &\ \ =\ \ & 
      \J{[f => [x => x]]} 
      \\
      \textbf{s} 
      &\ \ =\ \ & 
      \J{[n => [f => [x => f(n(f(x)))]]]} 
      \\
      \textbf{n3} 
      &\ \ =\ \ & 
      \textbf{s}\J{(}\textbf{s}\J{(}\textbf{s}\J{(}\textbf{z}\J{)}\J{)}\J{)} 
    \end{array}
  \end{array}
  \]
\end{definition}

@@

\begin{definition} 
  \label{def:experimental_typing_context}
  \emph{experimental typing context}
  \tiny
  \nopad
  \[
  \begin{array}[t]{ll}
    \begin{array}[t]{r@{}c@{}l}
      \textbf{ctx} & = &  \hfill
      \\
      \J{head} &\J{ : }& \J{ALL[T] }\textbf{List}(\J{T})\J{ -> T}
      \\
      \J{tail} &\J{ : }& \J{ALL[T] }\textbf{List}(\J{T})\J{ -> }\textbf{List}(\J{T})
      \\
      \J{nil} &\J{ : }& \J{ALL[T] }\textbf{List}(\J{T})
      \\
      \J{single} &\J{ : }& \J{ALL[T] T -> }\textbf{List}(\J{T})
      \\
      \J{append} &\J{ : }& \J{ALL[T] }\textbf{List}(\J{T}) \J{ -> } \textbf{List}(\J{T}) \J{ -> } \textbf{List}(\J{T})
      \\
      \J{length} &\J{ : }& \J{ALL[T] }\textbf{List}(\J{T}) \J{ -> } \textbf{Nat}
      \\
      \J{const} &\J{ : }& \J{ALL[T U] T -> U -> T}
      \\
      \J{id} &\J{ : }& \J{ALL[T] T -> T}
      \\
      \J{ids} &\J{ : }& \textbf{List}(\J{ALL[T] T -> T}\J{)}
      \\
      \J{inc} &\J{ : }& \textbf{Nat}\J{ -> }\textbf{Nat}
      \\
      \J{choose} &\J{ : }& \J{ALL[T] T -> T -> T}
      \\
      \J{poly} &\J{ : }& \J{(ALL[T] T -> T) -> } \textbf{Nat}\J{*}\textbf{Bool}
      \\
      \J{fst} &\J{ : }& \J{ALL[T U] T*U -> T}
      \\
      \J{sort} &\J{ : }& \J{ALL[T] (T * T -> } \textbf{Bool} \J{) -> }\textbf{List}(\J{T})\J{ -> }\textbf{List}(\J{T})
      \\
      \J{scalarCmp} &\J{ : }& \textbf{Nat} \J{\*} \textbf{Nat} \J{ -> } \textbf{Bool} 
      \\
      \J{lexicoCmp} &\J{ : }& \textbf{List}(\textbf{Nat}) \J{\*} \textbf{List}(\textbf{Nat}) \J{ -> } \textbf{Bool} 
    \end{array}
    &
    \begin{array}[t]{r@{}c@{}l}
      \\
      \J{auto} &\J{ : }& \J{(ALL[T] T -> T) -> (ALL[T] T -> T)}
      \\
      \J{auto'} &\J{ : }& \J{ALL[U] (ALL[T] T -> T) -> U -> U}
      \\
      \J{map} &\J{ : }& \J{ALL[T U] (T -> U) -> } \textbf{List}(\J{T}) \J{ -> } \textbf{List}(\J{U})
      \\
      \J{app} &\J{ : }& \J{ALL[T U] (T -> U) -> T -> U}
      \\
      \J{revapp} &\J{ : }& \J{ALL[T U] T -> (T -> U) -> U}
      \\
      \J{runST} &\J{ : }& \J{ALL[T] (ALL[U] } \textbf{ST}(\J{U}, \J{T})\J{) -> T}
      \\
      \J{argST} &\J{ : }& \J{ALL[U] } \textbf{ST}(\J{U}, \textbf{Nat})
      \\
      \J{zero} &\J{ : }& \textbf{ChurchNum}
      \\
      \J{succ} &\J{ : }& \textbf{ChurchNum} \J{ -> } \textbf{ChurchNum}
      \\
      \J{foo} &\J{ : }& \J{ALL[T] (T -> T) -> } \textbf{List}(\J{T}) \J{ -> T}
      \\
      \J{g} &\J{ : }& \J{ALL[T] }\textbf{List}(\J{T}) \J{ -> } \textbf{List}(\J{T}) \J{ -> T}
      \\
      \J{k} &\J{ : }& \J{ALL[T] T -> } \textbf{List}(\J{T}) \J{ -> T}
      \\
      \J{h} &\J{ : }& \textbf{Nat} \J{-> ALL[T] T -> T}
      \\
      \J{l} &\J{ : }& \textbf{List}(\J{ALL[T] } \textbf{Nat} \J{ -> T -> T}\J{)}
      \\
      \J{r} &\J{ : }& \J{(ALL[T] T -> ALL[U] B -> B) -> }\textbf{Nat}
    \end{array}
  \end{array}
  \]
\end{definition}

\noindent
Additionally, for the purpose of experiment, it is useful to assume that   
the typing of some identifiers have already been learned or verified.

\begin{experiment}
  \label{exp:part_1}
  Something
  \hfill
  \scriptsize
  \nopad
  \begin{center}
  \begin{tabular}{l m{30em} >{\centering}m{5em} >{\centering}m{5em} >{\centering\arraybackslash}m{5em} } 
    \multicolumn{5}{l}{
      \small
      Does the implementation admit an expression under typing constraints \textbf{ctx}?
    } \\

@@

    \sbottomrule
  \end{tabular}
  \end{center}
\end{experiment}

\noindent
In the experiment on functions, we 
are interested to observe if our implementation (\textbf{Extrinsic})
can admit various programs on functions. 
For our implementation, we reject a program if the implementation
finds a contradiction, diverges, or infers the top type. 
In our system, we infer a top type when the top level 
expression represents a function and all of its branches 
have errors. Since the top level function is not applied,
the program technically cannot produce an error, but it's 
perfectly useless and would produce an error if applied to any value. 
We compare our system to SuperF and Super without Distributivity (\textbf{SF no D}).

@@

\begin{experiment}
  \label{experi:part_2}
  \hfill
  \scriptsize
  \nopad
  \begin{center}
  \begin{tabular}{l m{30em} >{\centering}m{5em} >{\centering}m{5em} >{\centering\arraybackslash}m{5em} } 
    \multicolumn{5}{l}{
      \small
      Does the implementation admit an expression under typing constraints \textbf{ctx}?
    } \\

@@

    \sbottomrule
  \end{tabular}
  \end{center}
\end{experiment}

\begin{experiment}
  \label{experi:part_3}
  \hfill
  \scriptsize
  \nopad
  \begin{center}
  \begin{tabular}{l m{30em} >{\centering}m{5em} >{\centering}m{5em} >{\centering\arraybackslash}m{5em} } 
    \multicolumn{5}{l}{
      \small
      Does the implementation admit an expression under typing constraints \textbf{ctx}?
    } \\

@@

    \sbottomrule
  \end{tabular}
  \end{center}
\end{experiment}

\begin{experiment}
  \label{experi:part_4}
  \hfill
  \nopad
  \scriptsize
  \begin{center}
  \begin{tabular}{l m{30em} >{\centering}m{5em} >{\centering}m{5em} >{\centering\arraybackslash}m{5em} } 
    \multicolumn{5}{l}{
      \small
      Does the implementation admit an expression under typing constraints \textbf{ctx}?
    } \\

@@

  \end{tabular}
  \end{center}
\end{experiment}

\begin{experiment}
  \label{experi:part_5}
  \hfill
  \nopad
  \scriptsize
  \begin{center}
  \begin{tabular}{p{1em} l >{\centering}p{5em} >{\centering\arraybackslash}p{5em}} 
    \multicolumn{4}{l}{
      \small
      Does the implementation admit an expression under typing constraints \textbf{ctx}?
    } \\

@@

  \end{tabular}
  \end{center}
\end{experiment}

@@

\noindent
The experiment on programs with constructors demonstrate
that our system is roughly as expressive as untyped languages
without sacrificing safety guarantees.
Our system can easily mix structures together, reducing
the need for programmers to call different interfaces for different data.  
If there is an error in one branch of a function, the
system can still admit the program by ensuring 
that the erroneous branch is never used, and there
is at least one branch that could be used.

>>>>

\newtheorem{instances}{Instances}[section]

@@

      \delta \satisfies e \hastype \J{EXI[}\Theta\J{]}\Delta \J{:} \tau
      &\iff& 
      \exists \delta' \qua
      \textbf{dom}(\delta') \subseteq \Theta
      \up
      \delta' \oplus \delta \satisfies \Delta 
      \up
      \delta' \oplus \delta \satisfies e \hastype \tau

@@

      \delta \satisfies e \hastype \J{ALL[}\Theta\J{]}\Delta \J{:} \tau
      &\iff& 
      \forall \delta' \qua
      \textbf{dom}(\delta') \subseteq \Theta
      \implies
      \delta' \oplus \delta \satisfies \Delta 
      \implies
      \delta' \oplus \delta \satisfies e \hastype \tau

@@

      \delta \satisfies e \hastype \alpha 
      &\iff& 
      \VDash e \hastype \delta(\alpha) 
  \end{array}
  \]
\end{definition}

@@

    \inferrule {
      \alpha \neq \alpha_l
      \\\\
      \textbf{invert}(\alpha \entails \Delta) = \Delta' 
      \\
      \textbf{pack}^-(\alpha\ \alpha_l\ \Theta^\dagger \entails \Theta, \Delta', \alpha_l \J{*} \tau_r) = \tau'
      \\\\
      \textbf{factor}(\alpha \entails \tau' \cdot \J{left}) = \tau_l
      \\
      \textbf{factor}(\alpha \entails \tau' \cdot \J{right}) = \tau_r'
      \\\\
      \alpha_l \wr^+ \tau_r'
      \\
      % founded upper bound
      \alpha \downarrow \tau_l \subtypes \tau_l'
      \\
      \tau_r'[\alpha_l\slash \J{LFP[} \alpha \J{]} \tau_l'] = \tau_r''
    } {
      \Theta^\dagger \entails \alpha \circlearrowleft (\Theta, \Delta, \alpha_l \J{->}\tau_r) \subtypes \alpha_l \J{->} \J{LFP[} \alpha \J{]} \tau_r''
    }
  \end{mathpar}
\end{definition}

@@

\begin{definition} 
  \label{def:proof_subtyping_difference_intro}
  \emph{proof subtyping (difference introduction)}
  \hfill
  \small
  \boxed{\Theta, \Delta \entails \tau \subtypes \tau_l \typdiff \tau_r \given \Theta', \Delta'}
  \nopad
  \begin{mathpar}
    \inferrule {
      \TODO{Update proof, deriving that lower does not intersect with subtraction}
      \\\\
      \TODO{in lean, prove that subtyping is equivalent to decidable definition under restricted subtyping form}
      \\\\
      \TODO{the pattern restriction on both sides is more restrictive than subtyping restriction}
      \\\\
      \epsilon \pattype \tau_r
      \\
      \Theta, \Delta \not\dentails \tau \subtypes \tau_r
      \\
      \Theta, \Delta \not\dentails \tau_r \subtypes \tau
      \\
      \Theta, \Delta \entails \tau \subtypes \tau_l \given \Theta', \Delta' 
    } {
      \Theta, \Delta \entails 
      \tau \subtypes \tau_l \typdiff \tau_r
      \given \Theta', \Delta'
    }

    \inferrule {
      \TODO{proof rules should not require induction to construct premises}
      \\\\
      \epsilon \pattype \tau_r
      \\
      \alpha \prec \tau
      \\
      \alpha \wr^+ \tau
      \\\\
      \Theta, \Delta \not\dentails \tau[\alpha\slash\J{BOT}]^1 \subtypes \tau_r
      \\
      \Theta, \Delta \not\dentails \tau_r \subtypes \tau[\alpha\slash\J{BOT}]^{\textbf{height}(\tau_r)}
      \\
      \Theta, \Delta \entails \tau \subtypes \tau_l \given \Theta', \Delta' 
    } {
      \Theta, \Delta \entails 
      \J{LFP[} \alpha \J{]} \tau \subtypes \tau_l \typdiff \tau_r
      \given \Theta', \Delta'
    }

    \inferrule {
      \TODO{meta-theory}
      \\\\
      \epsilon \pattype \tau_r
      \\
      \alpha \prec \tau % struct-less-than
      \\
      \alpha \wr^+ \tau
      \\
    } {
      \Theta, \Delta \not\dentails \tau[\alpha\slash\J{BOT}]^1 \subtypes \tau_r
      \implies
      \nexists n \qua \Theta, \Delta \dentails \tau[\alpha\slash\J{BOT}]^n \subtypes \tau_r
    }

    \inferrule {
      \TODO{meta-theory}
      \\\\
      \epsilon \pattype \tau_r
      \\
      \alpha \prec \tau
      \\
      \alpha \wr^+ \tau
      \\
    } {
      \Theta, \Delta \not\dentails \tau_r \subtypes \tau[\alpha\slash\J{BOT}]^{\textbf{height}(\tau_r)}
      \implies
      \nexists n \qua \Theta, \Delta \dentails \tau_r \subtypes \tau[\alpha\slash\J{BOT}]^n
    }
  \end{mathpar}
\end{definition}

@@

\section{General Theory}
\label{sec:general_theory}

@@

\section{Instance Theory}
\label{sec:}

\noindent
From the declarative definitions, we extracted procedures to infer types and 
solve subtyping.
We constructed tactics to automatically prove 
static typing on instances of expressions and their inferred types.

We constructed and proved instances with both functions and constructors. 
These two categories have historically been tackled independently of each other, where parametric types
tend to rely on notions of instantiation, and structural types rely on notions of equality or subtyping.
Our system unifies these historically distinct problems by describing both kinds of types in terms of subtyping.
However, the distinction is useful for comparing our design to previous work.

We implemented an extrinsic type inference system that corresponds to the predicate definitions
presented thus far. 
To observe the expressiveness of our system on functions, 
we evaluated our implementation on the same suite of programs 
that SuperF was evaluated on, as presented in the paper on SuperF \cite{}.
Based on these instances, our system admits all reported be admitted by SuperF \cite{}, as well as some programs 
that SuperF is unable to admit.
Thus, our system appears to be more expressive on functions than the previous most expressive
type inference system that we know of.
To observe the expressiveness of our system on constructors, 
we constructed simple programs that represent the kind of flexible and reusable code
often found programs written in untyped scripting languages. 
We ran these examples in our implementation, and created analogous version to run
in the online implementation of MLScript \cite{},
which represents the theoretical system MLStruct \cite{}. 
All of these examples are admitted by our system, while MLStruct
can either requires intrinsic specifications to admit the program, 
or simply can't admit the program.

\begin{definition} 
  \label{def:expressions_and_type_instances}
  \emph{expression and type instances} 
  \scriptsize
  \nopad
  \[
  \begin{array}[t]{ll}
    \begin{array}[t]{r@{}c@{}l}
      \textbf{Bool} 
      &\ \ =\ \ & 
      \J{(<true>@ | <false>@)} 
      \\
      \textbf{Nat} 
      &\ \ =\ \ & 
      \J{(LFP[R] <zero>@ | <succ>R)} 
      \\
      \textbf{Even} 
      &\ \ =\ \ & 
      \J{(LFP[R] <zero>@ | <succ><succ>R)} 
      \\
      \textbf{List}(\alpha) 
      &\ \ =\ \ & 
      \J{(LFP[R] <nil>@ | <cons>((}\alpha\J{)*R))} 
      \\
      \textbf{ChurchNum} 
      &\ \ =\ \ & 
      \J{(ALL[T] (T -> T) -> T -> T)}
    \end{array}
    &
    \begin{array}[t]{r@{}c@{}l}
      \textbf{ST}(\alpha^\dagger, \alpha) 
      &\ \ =\ \ & 
      \J{(<tag>(}\alpha^\dagger \J{) <result>(} \alpha \J{))}
      \\
      \textbf{z} 
      &\ \ =\ \ & 
      \J{[f => [x => x]]} 
      \\
      \textbf{s} 
      &\ \ =\ \ & 
      \J{[n => [f => [x => f(n(f(x)))]]]} 
      \\
      \textbf{n3} 
      &\ \ =\ \ & 
      \textbf{s}\J{(}\textbf{s}\J{(}\textbf{s}\J{(}\textbf{z}\J{)}\J{)}\J{)} 
    \end{array}
  \end{array}
  \]
\end{definition}

@@

\begin{definition} 
  \label{def:tying_context_instance}
  \emph{tying context instance}
  \tiny
  \nopad
  \[
  \begin{array}[t]{ll}
    \begin{array}[t]{r@{}c@{}l}
      \textbf{ctx} & = &  \hfill
      \\
      \J{head} &\J{ : }& \J{ALL[T] }\textbf{List}(\J{T})\J{ -> T}
      \\
      \J{tail} &\J{ : }& \J{ALL[T] }\textbf{List}(\J{T})\J{ -> }\textbf{List}(\J{T})
      \\
      \J{nil} &\J{ : }& \J{ALL[T] }\textbf{List}(\J{T})
      \\
      \J{single} &\J{ : }& \J{ALL[T] T -> }\textbf{List}(\J{T})
      \\
      \J{append} &\J{ : }& \J{ALL[T] }\textbf{List}(\J{T}) \J{ -> } \textbf{List}(\J{T}) \J{ -> } \textbf{List}(\J{T})
      \\
      \J{length} &\J{ : }& \J{ALL[T] }\textbf{List}(\J{T}) \J{ -> } \textbf{Nat}
      \\
      \J{const} &\J{ : }& \J{ALL[T U] T -> U -> T}
      \\
      \J{id} &\J{ : }& \J{ALL[T] T -> T}
      \\
      \J{ids} &\J{ : }& \textbf{List}(\J{ALL[T] T -> T}\J{)}
      \\
      \J{inc} &\J{ : }& \textbf{Nat}\J{ -> }\textbf{Nat}
      \\
      \J{choose} &\J{ : }& \J{ALL[T] T -> T -> T}
      \\
      \J{poly} &\J{ : }& \J{(ALL[T] T -> T) -> } \textbf{Nat}\J{*}\textbf{Bool}
      \\
      \J{fst} &\J{ : }& \J{ALL[T U] T*U -> T}
      \\
      \J{sort} &\J{ : }& \J{ALL[T] (T * T -> } \textbf{Bool} \J{) -> }\textbf{List}(\J{T})\J{ -> }\textbf{List}(\J{T})
      \\
      \J{scalarCmp} &\J{ : }& \textbf{Nat} \J{\*} \textbf{Nat} \J{ -> } \textbf{Bool} 
      \\
      \J{lexicoCmp} &\J{ : }& \textbf{List}(\textbf{Nat}) \J{\*} \textbf{List}(\textbf{Nat}) \J{ -> } \textbf{Bool} 
    \end{array}
    &
    \begin{array}[t]{r@{}c@{}l}
      \\
      \J{auto} &\J{ : }& \J{(ALL[T] T -> T) -> (ALL[T] T -> T)}
      \\
      \J{auto'} &\J{ : }& \J{ALL[U] (ALL[T] T -> T) -> U -> U}
      \\
      \J{map} &\J{ : }& \J{ALL[T U] (T -> U) -> } \textbf{List}(\J{T}) \J{ -> } \textbf{List}(\J{U})
      \\
      \J{app} &\J{ : }& \J{ALL[T U] (T -> U) -> T -> U}
      \\
      \J{revapp} &\J{ : }& \J{ALL[T U] T -> (T -> U) -> U}
      \\
      \J{runST} &\J{ : }& \J{ALL[T] (ALL[U] } \textbf{ST}(\J{U}, \J{T})\J{) -> T}
      \\
      \J{argST} &\J{ : }& \J{ALL[U] } \textbf{ST}(\J{U}, \textbf{Nat})
      \\
      \J{zero} &\J{ : }& \textbf{ChurchNum}
      \\
      \J{succ} &\J{ : }& \textbf{ChurchNum} \J{ -> } \textbf{ChurchNum}
      \\
      \J{foo} &\J{ : }& \J{ALL[T] (T -> T) -> } \textbf{List}(\J{T}) \J{ -> T}
      \\
      \J{g} &\J{ : }& \J{ALL[T] }\textbf{List}(\J{T}) \J{ -> } \textbf{List}(\J{T}) \J{ -> T}
      \\
      \J{k} &\J{ : }& \J{ALL[T] T -> } \textbf{List}(\J{T}) \J{ -> T}
      \\
      \J{h} &\J{ : }& \textbf{Nat} \J{-> ALL[T] T -> T}
      \\
      \J{l} &\J{ : }& \textbf{List}(\J{ALL[T] } \textbf{Nat} \J{ -> T -> T}\J{)}
      \\
      \J{r} &\J{ : }& \J{(ALL[T] T -> ALL[U] B -> B) -> }\textbf{Nat}
    \end{array}
  \end{array}
  \]
\end{definition}

\noindent
Additionally, it is illustrative to assume that   
the typing of some identifiers have already been learned or verified.

\begin{instances}
  \label{exp:part_1}
  Something
  \hfill
  \scriptsize
  \nopad
  \begin{center}
  \begin{tabular}{l m{30em} >{\centering}m{5em} >{\centering}m{5em} >{\centering\arraybackslash}m{5em} } 
    \multicolumn{5}{l}{
      \small
      Does the implementation admit an expression under typing constraints \textbf{ctx}?
    } \\

@@

    \sbottomrule
  \end{tabular}
  \end{center}
\end{instances}

\noindent
In the instances with functions, we 
are interested to observe if our implementation (\textbf{Extrinsic})
can admit various programs on functions. 
For our implementation, we reject a program if the implementation
finds a contradiction, diverges, or infers the top type. 
In our system, we infer a top type when the top level 
expression represents a function and all of its branches 
have errors. Since the top level function is not applied,
the program technically cannot produce an error, but it's 
perfectly useless and would produce an error if applied to any value. 
We compare our system to SuperF and Super without Distributivity (\textbf{SF no D}).

@@

\begin{instances}
  \label{experi:part_2}
  \hfill
  \scriptsize
  \nopad
  \begin{center}
  \begin{tabular}{l m{30em} >{\centering}m{5em} >{\centering}m{5em} >{\centering\arraybackslash}m{5em} } 
    \multicolumn{5}{l}{
      \small
      Does the implementation admit an expression under typing constraints \textbf{ctx}?
    } \\

@@

    \sbottomrule
  \end{tabular}
  \end{center}
\end{instances}

\begin{instances}
  \label{experi:part_3}
  \hfill
  \scriptsize
  \nopad
  \begin{center}
  \begin{tabular}{l m{30em} >{\centering}m{5em} >{\centering}m{5em} >{\centering\arraybackslash}m{5em} } 
    \multicolumn{5}{l}{
      \small
      Does the implementation admit an expression under typing constraints \textbf{ctx}?
    } \\

@@

    \sbottomrule
  \end{tabular}
  \end{center}
\end{instances}

\begin{instances}
  \label{experi:part_4}
  \hfill
  \nopad
  \scriptsize
  \begin{center}
  \begin{tabular}{l m{30em} >{\centering}m{5em} >{\centering}m{5em} >{\centering\arraybackslash}m{5em} } 
    \multicolumn{5}{l}{
      \small
      Does the implementation admit an expression under typing constraints \textbf{ctx}?
    } \\

@@

  \end{tabular}
  \end{center}
\end{instances}

\begin{instances}
  \label{experi:part_5}
  \hfill
  \nopad
  \scriptsize
  \begin{center}
  \begin{tabular}{p{1em} l >{\centering}p{5em} >{\centering\arraybackslash}p{5em}} 
    \multicolumn{4}{l}{
      \small
      Does the implementation admit an expression under typing constraints \textbf{ctx}?
    } \\

@@

  \end{tabular}
  \end{center}
\end{instances}

@@

\noindent
The instances with constructors demonstrate
that our system is roughly as expressive as untyped languages
without sacrificing safety guarantees.
Our system can easily mix structures together, reducing
the need for programmers to call different interfaces for different data.  
If there is an error in one branch of a function, the
system can still admit the program by ensuring 
that the erroneous branch is never used, and there
is at least one branch that could be used.
